Fractional Decomposition Tree Algorithm: A tool for studying the integrality gap of Integer Programs

### 2022-05-05 FDT2020/conclusion.tex
comments almost all done
Added some disclaimers, etc, to the first paragraph of the experimental section

Our experiments in Section~\ref{sec:experiment} give a proof of concept. They show that its plausible FDT will have practical benefit as an IP heuristic for problems with appropriate structure. FDT performance will likely improve in a future more high-performance version coded in C or C++, perhaps able to take advantage of the low-level parallelism available on all modern platforms (even laptops).  This will allow tests on larger instances. We leave as future work a more comprehensive set of experiments to determine on which kinds of problems FDT is likely to outperform other general heuristics. Any consistent structure of such instances could lead to proofs of better approximation bounds.

>>>>

Our experiments in Section~\ref{sec:experiment} give a proof of concept. They show that it's plausible FDT will have practical benefit as an IP heuristic for problems with appropriate structure. FDT performance will likely improve in a future more high-performance version coded in C or C++, perhaps able to take advantage of the low-level parallelism available on all modern platforms (even laptops).  This will allow tests on larger instances. We leave as future work a more comprehensive set of experiments to determine on which kinds of problems FDT is likely to outperform other general heuristics. Any consistent structure of such instances could lead to proofs of better approximation bounds. Furthermore, in FDT we solve somewhat similar LPs iteratively. Hence, the running time of the algorithm can benefit from warm-starting the solver for the LPC problem (see \cite{warmstart} for a successful application of this approach.)

### 2022-05-05 FDT2020/main-section.tex
Added some disclaimers, etc, to the first paragraph of the experimental section

The FDT algorithm for 2EC iteratively applies Lemmas \ref{LPC2EC} and \ref{2ecpruning} to variables $x_1,\ldots,x_t$ to obtain leaf point solutions $L_t$. Finally, we just need to apply Lemma \ref{rounddown} to obtain the 2-edge-connected multi-subgraphs from every solution in $L_t$. Since $x$ is an extreme point we have $t\leq 2|V|-1$ \cite{boydpulley}. By Lemma \ref{2ecpruning} we have
\begin{align*}
\sum_{(x,\lambda)\in L_t} \frac{\lambda}{\sum_{(x,\lambda)\in L_t}\lambda} \floor{x} \leq \frac{1}{\sum_{(x,\lambda)\in L_t}\lambda} \sum_{(x,\lambda)\in L_t} \lambda {x} \leq g^t_{\2ec} x^*.
\end{align*}
\section{Computational Experiments with FDT}\label{sec:experiment}
We ran FDT on three network design problems: VC, TAP and 2EC.

@@

As described in Section~\ref{sec:2EC-intro}, when the LP optimum for (\ref{eq:subtour}) is a Carr-Vempala point (see Figure~\ref{fig:CVpoint}), the integrality gap of that instance is equal to that of 2EC. So in that sense, these are the hardest fractional points to round to a feasible solution. We now compare the quality of FDT's solution from a Carr-Vempala point to BOMC. BOMC (and simply Christofides) provide the best known upper bound on $g(\2ec)$\footnote{Recently, \cite{KKO} improved over Christofides algorithm by a small factor.}. Let us remind our reader the Christofides algorithm for TSP: given a graph with edge costs find a minimum spanning tree of the graph.  
%Then, since TSP runs on a complete graph, add a perfect matching on the vertices that have odd degree in the spanning tree to make the (multi)-subgraph 2-edge-connected. If we stop that algorithm without shortcutting, we have a feasible solution for Metric-2EC (given a complete graph, find the minimum-cost 2-edge-connected subgraph, with no multiedges).\arash{The following sentence is not accurate, Metric-TSP is the same as TSP on a non-complete graph where we can take multiple copies of the edges, Metric-2EC (given complete graph with metric costs find the minimum cost 2-edge-connected SUBgraph -no doubling of edges-) is the same as 2EC on a non-complete graph where we can take multiple copies of edges.}\cindy{It sounded like 2EC doesn't assume a complete graph.  But I removed the sentence.  Is the following OK?}\arash{Still not quite accurate. The sentence gives the impression that changing matching with $T$-join is the difference between TSP and 2EC, but its the difference between working with complete graph with metric costs and a non-complete graph where doubling edges is allowed.} 
Next, we add an $O$-join, where $O$ is the set of odd-degree vertices in the spanning tree~\footnote{For graph $G=(V,E)$ and $O\subseteq V$ with $|O|$ even, an $O$-join of $G$ is a subgraph of $G$ that has odd degree on the vertices in $O$ and even degree on vertices in $V\setminus O$. Graphs always have an even number of odd vertices since every edge has two endpoints, giving an even number of endpoints.}. Note that the solutions to the TSP are also 2-edge-connected, which means that Christofides algorithm also provides solutions to the 2EC.
The best-of-many Christofides (BOMC) algorithm~\cite{AKS15} works better than the classic Christofides algorithm in practice~\cite{GW17}.  The BOMC algorithm samples many spanning trees from the LP relaxation (with edges chosen proportionally to the LP value), augments each spanning tree with minimum cost join on the odd degree vertices of the spanning tree, and takes the best solution.

>>>>

The FDT algorithm for 2EC iteratively applies Lemmas \ref{LPC2EC} and \ref{2ecpruning} to variables $x_1,\ldots,x_t$ to obtain leaf point solutions $L_t$. Finally, we just need to apply Lemma \ref{rounddown} to obtain the 2-edge-connected multi-subgraphs from every solution in $L_t$. Since $x$ is an extreme point we have $t\leq 2|V|-1$ \cite{boydpulley}. By Lemma \ref{2ecpruning} we have
\begin{align*}
\sum_{(x,\lambda)\in L_t} \frac{\lambda}{\sum_{(x,\lambda)\in L_t}\lambda} \floor{x} \leq \frac{1}{\sum_{(x,\lambda)\in L_t}\lambda} \sum_{(x,\lambda)\in L_t} \lambda {x} \leq g^t_{\2ec} x^*.
\end{align*}
\section{Computational Experiments with FDT}\label{sec:experiment}
We ran FDT on three network design problems: VC, TAP and 2EC. These experiments, especially those for 2EC, illustrate possible uses of FDT to study the integrality gap of a problem. We also compare FDT the original feasibility pump algorithm~\cite{fp1}\footnote{We coud not find code for feasibility pump 2.0~\cite{fp2}.} or problem-specific approximation algorithms. Our intent is to show that FDT's performance in practice is far better than the provable bound on integrality gap and the run time is reasonable. We are not recommending FDT as a replacement for other general heuristics, but it may have a place in a suite of heuristics for some problems.

@@

As described in Section~\ref{sec:2EC-intro}, when the LP optimum for (\ref{eq:subtour}) is a Carr-Vempala point (see Figure~\ref{fig:CVpoint}), the integrality gap of that instance is equal to that of 2EC. So in that sense, these are the hardest fractional points to round to a feasible solution. We now compare the quality of FDT's solution from a Carr-Vempala point to BOMC. BOMC (and simply Christofides) provide the best known upper bound on $g(\2ec)$\footnote{Recently, \cite{KKO} gave an approximation algorithm that improves upon Christofides' bound by a small factor.}. Let us remind our reader the Christofides algorithm for TSP: given a graph with edge costs find a minimum spanning tree of the graph.  
%Then, since TSP runs on a complete graph, add a perfect matching on the vertices that have odd degree in the spanning tree to make the (multi)-subgraph 2-edge-connected. If we stop that algorithm without shortcutting, we have a feasible solution for Metric-2EC (given a complete graph, find the minimum-cost 2-edge-connected subgraph, with no multiedges).\arash{The following sentence is not accurate, Metric-TSP is the same as TSP on a non-complete graph where we can take multiple copies of the edges, Metric-2EC (given complete graph with metric costs find the minimum cost 2-edge-connected SUBgraph -no doubling of edges-) is the same as 2EC on a non-complete graph where we can take multiple copies of edges.}\cindy{It sounded like 2EC doesn't assume a complete graph.  But I removed the sentence.  Is the following OK?}\arash{Still not quite accurate. The sentence gives the impression that changing matching with $T$-join is the difference between TSP and 2EC, but its the difference between working with complete graph with metric costs and a non-complete graph where doubling edges is allowed.} 
Next, we add an $O$-join, where $O$ is the set of odd-degree vertices in the spanning tree~\footnote{For graph $G=(V,E)$ and $O\subseteq V$ with $|O|$ even, an $O$-join of $G$ is a subgraph of $G$ that has odd degree on the vertices in $O$ and even degree on vertices in $V\setminus O$. Graphs always have an even number of odd vertices since every edge has two endpoints, giving an even number of endpoints.}. Note that the solutions to the TSP are also 2-edge-connected, which means that Christofides algorithm also provides solutions to the 2EC.
The best-of-many Christofides (BOMC) algorithm~\cite{AKS15} works better than the classic Christofides algorithm in practice~\cite{GW17}.  The BOMC algorithm samples many spanning trees from the LP relaxation (with edges chosen proportionally to the LP value), augments each spanning tree with minimum cost join on the odd degree vertices of the spanning tree, and takes the best solution.
